Applying Deep Learning to Acoustic Analysis in the PICO-60 Dark Matter Detector

### 2018-08-11 documentation/paper.tex
Removed highly misleading graph
Rearranged wording of introduction

In experiments for dark matter detection, one of the most important and challenging hurdles to overcome is that of background radiation. Unwanted particles from a variety of sources can produce event signatures that appear very similar to those of dark matter candidates.

@@

The objective of this study is to explore a potential methodology that resolves both of these problems using techniques based on machine learning. These include supervised learning algorithms that learn to separate calibration data sets, and semi-supervised learning algorithms that simultaneously learn an effective discriminator and improve the purity of their own training data sets.

@@

@@

The performance of supervised learning neural networks was evaluated using 2 metrics. One is a simple classification accuracy value, which corresponds to the number of validation examples the network classifies incorrectly. This effectively summarizes how often the network is outright correct or incorrect, but fails to capture how decisive it is.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{fourier_mlp_no_correction}
    \caption{\label{fourier_mlp_no_correction} Accuracy of $FourierMLP(\beta_{8})$ compared to Acoustic Parameter at discriminating between run types (98\% accuracy).}
\end{figure}

@@

One of the main purposes of semi-supervised learning is to handle the impure training data effectively, and learn to replicate the imperfectly represented underlying pattern (nuclear recoils versus alpha particles) rather than merely fitting to separate run types. Therefore, performance is quantified by how accurately a system is able to replicate Acoustic Parameter, rather than simply reproduce the labels provided.

>>>>

In the PICO-60 experiment, and in most other experiments for dark matter detection, one of the most important and challenging hurdles to overcome is that of background radiation. Unwanted particles from a variety of sources can produce event signatures that appear very similar to those of dark matter candidates.

@@

The objective of this study is to investigate the potential for machine learning techniques to address both of these issues. More specifically, in the context of the PICO-60 experiment, there are 2 objectives:

\begin{enumerate}
    \item Determine whether machine learning techniques can function comparably to a conventionally derived discriminator. Furthermore, develop several neural network architectures and apply them to different input formats, in order to find which are the most effective.

    \item Determine whether these techniques can be extended to develop an effective discriminator function based on incomplete and inaccurate initial information.

    For clarity, and with specific reference to the PICO-60 experiment, the intent is to accurately separate particle types using only calibration data for training. While calibration runs are known to be mostly correlated with particle types, the relatively large percentage of impurities has the potential to hinder conventional learning techniques.
\end{enumerate}

The benefit to experimental science would be to potentially allow the creation of a discriminator in cases where it would be difficult to construct one by conventional means. There may be various reasons for this:

\begin{enumerate}
    \item Limited ability to collect accurate calibration data, and/or imperfect correlation between calibration data and particle types.
    \item Unknown and/or complex interactions in the experimental apparatus make a conventional discriminator difficult to define, where machine learning techniques may be able to uncover hidden correlations in the input data.
\end{enumerate}

@@

Performance of machine learning systems was evaluated using 2 metrics. One is a simple classification accuracy value, which corresponds to the number of validation examples the network classifies incorrectly. This effectively summarizes how often the network is outright correct or incorrect relative to the baseline provided for it. However, this fails to capture how decisive the network is: whether its prediction is confidently high or low, or is only slightly closer to one edge of the spectrum.

@@

@@

@@
